---
title: "ggplot2"
output:
  ioslides_presentation:
    font_adjustment: -5
---
##
*Основные компоненты графика ggplot*:
<font size="4">

- ggplot() - *create a new ggplot*

- data - *data.frame format data*

- aes() - *construct aesthetic mappings*

- _layers_, _scales_, _coords_, _facets_, добавленные с помощью _+_
  - слои _layers_
  - геометрические фигуры _geoms_ (points, lines, polygons, etc.)
  - статистические преобразования _stats_ (binning and counting)

  - параметры _scales_ (colour, shape, size, legend and axes)

  - координаты _coords_

  - группировка _facets_

- ggsave() - *save a ggplot (or other grid object) with sensible defaults*

- qplot() quickplot() - *quick plot*

</font>

```{r message=F, warning = F, echo = F}
library(ggplot2)
```
##
*geom_col & geom_bar*
<font size="4">

- geom_col, geom_bar(stat="identity")
  - for data frame where one column represents the x position of each bar, and another column represents the vertical (y) height of each bar
- geom_bar(stat="identity") [by default]
  - for data has one row representing each case, and you want plot counts of the cases

</font>

##
*geom_col & geom_bar*
```{r fig.height=1.5, fig.width = 2, message=F, warning=F}
dat <- data.frame(
        time = factor(c("Lunch","Dinner"), levels=c("Lunch","Dinner")),
        total_bill = c(14.89, 17.23))
ggplot(data=dat, aes(x=time, y=total_bill)) + geom_col()
```

##
*geom_col & geom_bar*
```{r fig.height=1.5, fig.width =2, fontsize=3, message=F, warning=F}
ggplot(data=dat, aes(x=time, y=total_bill)) + geom_col()
ggplot(data=dat, aes(x=time, y=total_bill)) + geom_bar(stat="identity")
```

##
*geom_col*

```{r fig.height=1.5, fig.width = 2, message=F, warning=F}
ggplot() + geom_bar(data=dat, aes(x=time, y=total_bill), stat="identity")
```

```{r fig.height=1.5, fig.width = 2, message=F, warning = F}
ggplot() + geom_col(data=dat, aes(x=time, y=total_bill), fill = "white")
```

##
*geom_col*
```{r fig.height=2, fig.width = 4}
ggplot() +
         geom_col(data=dat,
                  aes(x=time, y=total_bill, fill = time),
                  colour = 'black')
```


##
*geom_col*

```{r echo = F, warning = F, message = F}
library(gcookbook)
library(dplyr)# Load gcookbook for the cabbage_exp data set

```


```{r fig.height=2}
cabbage_exp %>% head(3)
ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) +
        geom_col()
```

##
*geom_col*
```{r fig.height=1.5, fig.width = 3}
ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) +
        geom_col()
ggplot(cabbage_exp, aes(x = Date, y = Weight, fill = Cultivar)) +
        geom_col(position = "dodge")
```


##
*geom_bar*

```{r, echo=F, message = F, warning = F}
library(reshape2)
```

```{r, error=F, warning=F, fig.height=1.5}
tips %>% head(3)
ggplot(data=tips, aes(x=day)) + geom_bar()
```

```{r, echo=F, message = F, warning = F}
library(reshape2)
```




```{r, echo=F, message = F, warning = F}
library(reshape2)
```
```{r, echo=F}
library(reshape2)
```

##
*geom_boxplot*
```{r ,error=F, warning=F, message=F, fig.height=1.5, fig.width= 3}
ggplot(PlantGrowth, aes(x=group, y=weight)) +
        geom_boxplot()
ggplot(PlantGrowth, aes(x=group, y=weight, fill = group)) +
        geom_boxplot(colour="black")
```


##
*geom_path vs geom_line*
<font size=5>

- geom_path() - значения соединены в том порядке, в котором они представлены в df
- geom_line() - значения соединены в том порядке, в котором они представлены по оси x

</font>

```{r, echo=F}
recent <- economics[economics$date > as.Date("2013-01-01"), ]

```

##
*geom_path vs geom_line*
```{r echo = F, message = F, warning = F}
economics %>% head(3)
```
```{r fig.height=1.5, warning = F, message = F}
m <- ggplot(economics, aes(unemploy/pop, psavert))
m1 <- m + geom_path()
m2 <- m + geom_line()
```
```{r echo=F, message = F, warning=F, fig.height=1.8, fig.width = 4}
library(gridExtra)
grid.arrange(m1 + ggtitle('geom_path'), m2 + ggtitle('geom_line'), ncol = 2)
```


##
*geom_line vs geom_step*

```{r fig.height=1.8, message = F, warning = F}
recent %>% head(3)
g1 <- ggplot(recent, aes(date, unemploy)) + geom_line()
g2 <- ggplot(recent, aes(date, unemploy)) + geom_step()

```


```{r echo=F, fig.height=1.8}
library(gridExtra)
grid.arrange(g1 + ggtitle('geom_line'), g2 + ggtitle('geom_step'), ncol = 2)
```

##
*geom_line*

```{r fig.height=1.8}
c <- ggplot(economics, aes(x = date, y = pop))
c1 <- c + geom_line(colour = 'red', size = 20)
c2 <- c + geom_line(colour = 'blue', linetype = 2)
c3 <- c + geom_line(arrow = arrow(angle = 15, ends = "both", type = "closed"))

```

```{r echo=F, fig.height=1.8}
library(gridExtra)
grid.arrange(c1 + ggtitle('c1'), c2 + ggtitle('c2'), c3 + ggtitle('c3'), ncol = 3)
```


##
*geom_line*


```{r, fig.height=1.8}
x <- seq(0.01, .99, length.out = 100)
df <- data.frame(
  x = rep(x, 2),
  y = c(qlogis(x), 2 * qlogis(x)),
  group = rep(c("a","b"), each = 100)
)
ggplot(df, aes(x=x, y=y, group=group, fill=group)) + geom_line()

```


##
*geom_line + geom_point*
```{r echo = F, message = F, warning = F}
library(gridExtra)

```
```{r, error=F, warning=F, fig.height=3}
g <- ggplot(data=dat, aes(x=time, y=total_bill, group=1))
gg1 <- g + geom_line()
gg2 <- g + geom_line() + geom_point()
gg3 <- g + geom_line(colour="red", linetype="dashed", size=1.5) +
  geom_point(colour="red", size=4, shape=21, fill="white")
gg4 <- g + geom_line() + geom_point() +
        expand_limits(y=0) +
        xlab("Time of day") + ylab("Total bill") +
        ggtitle("Average bill for 2 people")
```

```{r fig.height = 1.8, echo = F, message = F, warning = F}
grid.arrange(gg1, gg2, gg3, gg4, ncol=4)

```


```{r, error=F, warning=F, message = F, echo = F}
set.seed(1234)
dat <- data.frame(cond = factor(rep(c("A","B"), each=200)),
                  rating = c(rnorm(200),rnorm(200, mean=.8)))

```

##
*geom_histogram, geom_density*

```{r fig.height=3}
g <- ggplot(dat, aes(x=rating))
gg1 <- g + geom_histogram()
gg2 <- g + geom_histogram(binwidth=.5, colour="black", fill="white")
gg3 <- g + geom_density()
```


```{r fig.height=1.5, echo = F}
grid.arrange(gg1, gg2, gg3, ncol=3)
```

##
*geom_histogram, geom_density*


```{r fig.height=3}
gg4 <- ggplot(dat, aes(x=rating)) +
        geom_density(alpha=.2, fill="#FF6666") +
        geom_histogram(binwidth=.5, colour="black", fill="white")
gg5 <- ggplot(dat, aes(x=rating)) +
        geom_histogram(aes(y=..density..), binwidth=.5,
                       colour="black", fill="white") +
        geom_density(alpha=.2, fill="#FF6666")
gg6 <- ggplot(dat, aes(x=rating)) +
        geom_density(fill="#FF6666") +
        geom_histogram(aes(y=..density..), alpha=.2,  binwidth=.5,
                       colour="black", fill="blue")


```


```{r fig.height=1.5, echo = F}
grid.arrange(gg4, gg5, gg6, ncol=3)
```



##
*geom_errorbar vs geom_pointrang*

```{r echo=F, error=F, warning=F, message=F}
library(dplyr)
library(gridExtra)

tg <- ToothGrowth
tgg <- group_by(tg, supp, dose)
tgs <- summarise(tgg,
                 len_mean = mean(len),
                 len_max = mean(len) + 1.96 * sd(len)/sqrt(length(len)),
                 len_min = mean(len) - 1.96 * sd(len)/sqrt(length(len))) %>% ungroup()
tgs %>% head(3)
```


```{r error=F, warning=F, message=F, fig.height=1.5}
g <- ggplot(tgs, aes(x=dose, y=len_mean, colour=supp))
gg1 <- g + geom_line() + geom_point() +
        geom_errorbar(aes(ymin=len_min, ymax=len_max), width=.1)
gg2 <- g + geom_line() +
         geom_pointrange(aes(ymin=len_min, ymax=len_max))
```

```{r, echo=F, fig.height=1.5}
grid.arrange(gg1, gg2, ncol = 2)
```


##
*geom_errorbar*

```{r , error=F, warning=F, message=F, fig.height=1.5, fig.width= 15}
gg1 <- ggplot(tgs, aes(x=dose, y=len_mean, colour=supp)) +
        geom_errorbar(aes(ymin=len_min, ymax=len_max), width=.1) + geom_line() + geom_point()
pd <- position_dodge(0.1) # move them .05 to the left and right
gg2 <- ggplot(tgs, aes(x=dose, y=len_mean, colour=supp)) +
        geom_errorbar(aes(ymin=len_min, ymax=len_max), width=.1, position=pd) + geom_line(position=pd) + geom_point(position=pd)

```

```{r, echo=F, fig.height=1.5}
grid.arrange(gg1, gg2, ncol = 2)
```


##
*geom_point + geom_smooth*
```{r echo=F, error=F, warning=F, message=F}
dat <- data.frame(cond = rep(c("A", "B"), each=10),
                  xvar = 1:20 + rnorm(20,sd=3),
                  yvar = 1:20 + rnorm(20,sd=3))

```


```{r, error=F, warning=F, message=F, fig.height=3, fig.width= 15}
head(dat)

gg1 <- ggplot(dat, aes(x=xvar, y=yvar)) + geom_point(shape=1)
gg2 <- ggplot(dat, aes(x=xvar, y=yvar)) + geom_point(shape=1) + geom_smooth(method=lm) # Add linear regression line (by default includes 95% confidence region)
gg3 <- ggplot(dat, aes(x=xvar, y=yvar)) + geom_point(shape=1) + geom_smooth(method=lm, se=FALSE) # Add linear regression line, don't add shaded confidence region
gg4 <- ggplot(dat, aes(x=xvar, y=yvar)) + geom_point(shape=1) + geom_smooth() # Add a loess smoothed fit curve with confidence region

```

```{r, echo=F,error=F, warning=F, message=F, fig.height=3, fig.width= 15}
grid.arrange(gg1, gg2, gg3, gg4, ncol = 4)
```


## *geom_text*, *geom_labels*
- label individual observations or groups of observations

```{r echo=F}
best_in_class <- mpg %>%
  group_by(class) %>%
  filter(row_number(desc(hwy)) == 1)

```

```{r, fig.height=3, fig.width= 15}
gg1 <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_text(aes(label = model), data = best_in_class)

gg2 <- ggplot(mpg, aes(displ, hwy)) +
  geom_point(aes(colour = class)) +
  geom_label(aes(label = model), data = best_in_class, nudge_y = 2, alpha = 0.5)
```

```{r, echo=F,error=F, warning=F, message=F, fig.height=3, fig.width= 15}
grid.arrange(gg1, gg2, ncol = 2)
```

## _scales_

*scale_[название переменной]_[тип переменной]*

 - scale_x_continuous
 - scale_y_continuous
 - scale_fill_discrete
 - scale_colour_discrete
 - scale_colour_hue
 - scale_fill_manual
 - scale_alpha_manual
 - scale_shape_manual

## _scales_

```{r}
bp <- ggplot(data=PlantGrowth, aes(x=group, y=weight, fill=group)) + geom_boxplot()
bp1 <- bp + scale_fill_discrete(guide=FALSE)
bp2 <- bp + scale_fill_discrete(guide = guide_legend(reverse=TRUE))
bp3 <- bp + scale_fill_discrete(breaks=c("trt1","ctrl","trt2"))
bp4 <- bp + scale_fill_discrete(labels=c("Control", "Treatment 1", "Treatment 2"))
bp5 <- bp + scale_fill_manual(labels=c("Control", "Treatment 1", "Treatment 2"),
                              values=c("#999999", "#E69F00", "#56B4E9"))

```

```{r, echo=F,error=F, warning=F, message=F, fig.height=3, fig.width = 18}
grid.arrange(bp1, bp2, bp3, bp4, bp5, ncol = 5)
```

## _Scales_
```{r, echo = F}
dat1 <- data.frame(
  sex = factor(c("Female","Female","Male","Male")),
  time = factor(c("Lunch","Dinner","Lunch","Dinner"), levels=c("Lunch","Dinner")),
  total_bill = c(13.53, 16.81, 16.24, 17.42)
)
head(dat1)
```


```{r, error=F, warning=F}
gg <- ggplot(data=dat1, aes(x=time, y=total_bill, group=sex, colour=sex)) + geom_line(aes(linetype=sex), size=1) + geom_point(aes(shape=sex), size=3, fill="white") + expand_limits(y=c(0,20))
gg1 <-  gg + scale_colour_hue(name="Sex of payer", l=50)

gg2 <- gg1 + scale_shape_manual(name="Sex of payer",
                                values=c(22,1)) +
  scale_linetype_discrete(name="Sex of payer") +
  xlab("Time of day") + ylab("Total bill") + # Set axis labels
  ggtitle("Average bill for 2 people") +     # Set title
  theme_bw() +
  theme(legend.position=c(.8, .4))           # Position legend inside
# This must go after theme_bw

```

```{r, echo=F,error=F, warning=F, message=F, fig.height=3, fig.width = 18}
grid.arrange(gg, gg1, gg2, ncol = 3)
```

## *scale*
```{r}
dsamp <- diamonds[sample(nrow(diamonds), 1000), ]
d <- ggplot(dsamp, aes(carat, price)) + geom_point(aes(colour = clarity))
dd1 <- d + scale_colour_hue(expression(clarity[beta]))
dd2 <- d + scale_colour_hue(l = 40, c = 30)
dd3 <- d + scale_colour_hue(l = 70, c = 30)
```

```{r, echo=F,error=F, warning=F, message=F, fig.height=3, fig.width = 18}
grid.arrange(d, dd1, dd2, dd3, ncol = 4)
```

## *scales*

```{r}
head(mpg)
gg <- ggplot(mpg, aes(displ, hwy)) + geom_point()
gg1 <- gg + scale_y_continuous(name = 'Highway fuel economy (mpg)')
gg2 <- gg + scale_y_continuous(breaks = seq(0, 40, by = 5))
gg3 <- gg + scale_x_continuous(limits = c(0, 10))
gg4 <- gg + scale_x_continuous(labels = scales::comma)
```

```{r, echo=F,error=F, warning=F, message=F, fig.height=3, fig.width = 15}
grid.arrange(gg1, gg2, gg3, gg4, ncol = 4)
```

## *labs*
- *title* - to add labels
- *subtitle* - to add additional detail in a smaller font beneath the title
- *caption* - to add text at the bottom right of the plot, often used to describe the source of the data
- to replace the axis and legend titles

# *labs*
```{r warning=F, error=F, message=F}
head(mpg)
gg <- ggplot(mpg, aes(displ, hwy)) + geom_point(aes(color = class)) + geom_smooth(se = FALSE)
gg1 <- gg + labs(title = "Fuel efficiency generally decreases with engine size",
                 subtitle = "Two seaters (sports cars) are an exception because of their light weight",
                 caption = "Data from fueleconomy.gov")
```

```{r, echo=F,error=F, warning=F, message=F, fig.height=3, fig.width = 18}
grid.arrange(gg, gg1, ncol = 2)
```

## *labs*

```{r warning=F, error=F, message=F}
head(mpg)
gg <- ggplot(mpg, aes(displ, hwy)) + geom_point(aes(color = class)) + geom_smooth(se = FALSE)
gg1 <- gg + labs(x = "Engine displacement (L)",
                 y = "Highway fuel economy (mpg)",
                 colour = "Car type")
```

```{r, echo=F,error=F, warning=F, message=F, fig.height=3, fig.width = 18}
grid.arrange(gg, gg1, ncol = 2)
```


## *Facetting*
- generates small multiples, each displaying a different subset of the data.

```{r}
head(tips)
sp <- ggplot(tips, aes(x=total_bill, y=tip/total_bill)) + geom_point(shape=1)
sp1 <- sp + facet_grid(. ~ sex)
sp2 <- sp + facet_grid(sex ~ day)
sp3 <- sp + facet_wrap( ~ day, ncol=2)
sp4 <- sp + facet_grid(sex ~ day) + theme(strip.text.x = element_text(size=8, angle=75),
                                          strip.text.y = element_text(size=12, face="bold"),
                                          strip.background = element_rect(colour="red", fill="#CCCCFF"))
```

```{r, echo=F,error=F, warning=F, message=F, fig.height=3, fig.width = 18}
grid.arrange(sp, sp1, sp2, sp3, sp4, ncol = 5)
```

